### 2020-03-01 ChapterBackground/chapterBackground.tex
Fixed: Corrections to Background

Real-time physics simulations (physics engines) are responsible for simulating the physical behaviour of objects and are based around Newtonian mechanics. The motion of the system is calculated given the forces acting on the system, this is known as the forward dynamics problem \cite{Boeing2007}.

@@

The numerical integrator used to update the positions of objects in the system, uses a discrete time value for each step, this is known as the time step. Each step simulates a set period of time. The shorter the time period simulated, the more physically accurate the simulation will be, however, each simulation step must take less time to execute than the time it is intending to simulate, otherwise the simulation would be unable to run in real-time. It is common for the time step to be a fixed value, though physics engines can support variable time steps.

@@

A Symplectic Euler (Semi-Implicit Euler) Integrator is often used in physics engines. Symplectic Euler is similar to explicit Euler, except that the updated velocity is used before calculating the new position:

@@

Collision response is the phase of the real-time physics loop which deals with integrating into the model the effect produced by contacts between entities within the system. Collision response will taken into account mass, coefficient of restitution (loss of kinetic energy during a collisions) and the location and angle of the collision. Collision response can be very simple to calculate for simple geometric shapes, but becomes expensive when more complex mesh representations are used.
%\subsection{Constraints}

@@

\subsubsection{Client-Server}
Client-Server architecture is a type of architecture in which the game is executed and game state is managed entirely by the server. Clients connect to the server and receive the necessary information about the game world. All messages and interactions are sent through the game server and the server is responsible for disseminating messages to the appropriate clients. The main challenge with this approach is scalability, as a single server can only support a limited number of players. This is normally solved with the addition of more servers, i.e. switching to a distributed server architecture. As all communication goes through the server, clients do not directly communicate with each other, meaning latency can be increased for interactions with other players.
As there is only one central version of the game state, consistency resolution is simplified to solving inconsistencies between server and client and the server treated as the correct master copy. In addition, fault-tolerance is simplified to keeping a single backup of the central server's sate, although this adds more complexity and cost, and can reduce scalability. Client-Server architectures are less prone to cheating than peer-to-peer based architectures, as the game provider has complete control over the game state \cite{P2PSurvey}.

@@

\subsubsection{Peer-to-Peer}
Peer-to-Peer, often referred to as`P2P', is a type of architecture in which the game is executed and game state is managed entirely by the peers, although peers are often only responsible for a subset of the game state, such as the entities they are currently interacting with. A peer-to-peer architecture is shown in Fig. \ref{fig_P2P}. Peers are all responsible for message forwarding to each other. In other words, peers act as bother server and client. Peer-to-Peer reduces the cost for the game provider of maintaining expensive servers and can reduce latency between players, as communication is performed directly between peers, rather than communicating through a server. In addition, it provides scalability in terms of compute power, as the more peers that join, the more resources available to the system. However, as more peers join, the greater the communication overhead is between peers. In addition, peer-to-peer is prone to the problems of fault-tolerance, data consistency and cheating by players. A peer is more likely to fail or experience unscheduled disconnection than a dedicated server. As each peer is responsible for the game state, it is possible for communication errors to occur, resulting in two peers with different game states or a peer maliciously altering the game state or the messages being sent to other peers in order to cheat. The fault-tolerance requirements of peer-to-peer can be reduced in short-session-based games, making these architectures better suited to those games \cite{P2PSurvey}. Peer-to-Peer architectures have received much research attention in previous years, with many solutions that attempt to address the problems above, through the use of techniques discussed in \cite{P2PSurvey}.

@@

\subsection{Reliability and Ordering layer}
RakNet uses UDP but also provides a reliability and ordering layer. When messages have the reliable option enabled, RakNet guarantees that UDP packets will arrive at their destination eventually. When messages. When messages have the ordered option enabled, RakNet guarantees that the packets are ordered at the destination. Ordering has the advantage of removing the need to handle out of order packets \cite{RakNetReliability}.

@@

In this section we discuss research into existing methods for distributing real-time physics across a cluster of nodes in a network.

@@

Distributed Real-Time Physics techniques fall into two main categories, those that use multiple instances of the same physics engine (as used by AP) and those that distribute particular aspects of the computational workload of a physics engine. We will now discuss both techniques.

@@

Techniques that distribute a particular aspect of the workload have not gained any commercial attention recently, unlike techniques that use multiple instances of the same physics engine, which we will now discuss.

@@

\section{Cloud Computing}
In the section a brief overview of cloud computing is given, deployment and service models are discussed. This study's implementation is deployed on the chosen cloud provider, Amazon Web Services (AWS). The latency and packet-loss conditions that affect AP are presented and these values will used to inform the correctness experiments performed on AP.

@@

In streamed gaming, each game instance resides on a single server. There is no technology to balance the real-time requirements of the game across multiple servers. The core problem is that all gaming technology is built and designed for single console/PC install and the greatest bottleneck is the inability to share physics calculations across machines. The research presented in this paper is intended for use in streamed gaming. The technique presented in this paper allows multiple servers to simulate physics, meaning physics simulations are no longer limited in size and complexity to that which can be achieved by only a single server. This allows for game streaming servers to make use of multiple physics servers, the numbers of which can grow and shrink as needed by the game and greatly increase the limit on size and complexity of physics within the game.

\section{Background Summary}
In physics simulations, there must be an object present in the solving phase for it to be considered in the overall solution of the scenario. A non-migratory approach requires a "ghost" representation of the object in a remote server to enable interaction. Given the calculations and discretisation steps, a "ghost" object takes up just as many resources in deriving a solution as a real object. This rules out the non-migratory approach for our problem. This leaves migratory.

Migratory approaches are concerned with managing the network traffic for those entities that could possibly exist on two servers but can only be solved on one. Such objects need to be placed with an owning server while minimising the effect of thrashing (where an algorithm frequently transfers objects between servers). In the rest of this paper, we describe our approach to solving this algorithmically, present how a working implementation was achieved, and present results evidencing our work. This is the first presentation of literature that can demonstrate real-time scalable server-side physics modelling and is a significant contribution to reducing the cost of commercialised streamed gaming.

>>>>

Real-time physics simulations (physics engines) are responsible for simulating the physical behaviour of objects and are based around Newtonian mechanics. The motion of the system is calculated given the forces acting on the system and this is known as the forward dynamics problem \cite{Boeing2007}.

@@

The numerical integrator used to update the positions of objects in the system, uses a discrete time value. Each discrete time value is known as a time step. Each step simulates a set period of time. The shorter the time period simulated, the more physically accurate the simulation will be, however, each simulation step must take less time to execute than the time it is intending to simulate, otherwise the simulation would be unable to run in real-time. It is common for the time step to be a fixed value, though physics engines can support variable time steps.

@@

A symplectic Euler (semi-implicit Euler) integrator is often used in physics engines. Symplectic Euler is similar to explicit Euler, except that the updated velocity is used before calculating the new position:

@@

Collision response is the phase of the real-time physics loop which deals with integrating into the model the effect produced by contacts between entities within the system. Collision response will take into account mass, coefficient of restitution (loss of kinetic energy during a collisions) and the location and angle of the collision. Collision response can be very simple to calculate for simple geometric shapes, but becomes expensive when more complex mesh representations are used.
%\subsection{Constraints}

@@

\subsubsection{Client-Server}
Client-Server architecture is a type of architecture in which the game is executed and game state is managed entirely by the server. Clients connect to the server and receive the necessary information about the game world. All messages and interactions are sent through the game server and the server is responsible for disseminating messages to the appropriate clients. The main challenge with this approach is scalability, as a single server can only support a limited number of players. This is normally solved with the addition of more servers, i.e. switching to a distributed server architecture. As all communication goes through the server, clients do not directly communicate with each other, meaning latency can be increased for interactions with other players.
As there is only one central version of the game state, consistency resolution is simplified to solving inconsistencies between server and client and the server treated as the correct master copy. In addition, fault-tolerance is simplified to keeping a single backup of the sate of the central server, although this adds more complexity and cost, and can reduce scalability. Client-Server architectures are less prone to cheating than peer-to-peer based architectures, as the game provider has complete control over the game state \cite{P2PSurvey}.

@@

\subsubsection{Peer-to-Peer}
Peer-to-Peer, often referred to as`P2P', is a type of architecture in which the game is executed and game state is managed entirely by the peers, although peers are often only responsible for a subset of the game state, such as the entities they are currently interacting with. A peer-to-peer architecture is shown in Fig. \ref{fig_P2P}. Peers are all responsible for message forwarding to each other. In other words, peers act as both server and client. Peer-to-Peer reduces the cost for the game provider of maintaining expensive servers and can reduce latency between players, as communication is performed directly between peers, rather than communicating through a server. In addition, it provides scalability in terms of compute power, as the more peers that join, the more resources available to the system. However, as more peers join, the greater the communication overhead is between peers. In addition, peer-to-peer is prone to the problems of fault-tolerance, data consistency and cheating by players. A peer is more likely to fail or experience unscheduled disconnection than a dedicated server. As each peer is responsible for the game state, it is possible for communication errors to occur, resulting in two peers with different game states or a peer maliciously altering the game state or the messages being sent to other peers in order to cheat. The fault-tolerance requirements of peer-to-peer can be reduced in short-session-based games, making these architectures better suited to those games \cite{P2PSurvey}. Peer-to-Peer architectures have received much research attention in previous years, with many solutions that attempt to address the problems above, through the use of techniques discussed in \cite{P2PSurvey}.

@@

\subsection{Reliability and Ordering layer}
RakNet uses UDP but also provides a reliability and ordering layer. When messages have the reliable option enabled, RakNet guarantees that UDP packets will arrive at their destination eventually. When messages have the ordered option enabled, RakNet guarantees that the packets are ordered at the destination. Ordering has the advantage of removing the need to handle out of order packets \cite{RakNetReliability}.

@@

In this section research into existing methods for distributing real-time physics across a cluster of nodes in a network will be discussed.

@@

Distributed Real-Time Physics techniques fall into two main categories, those that use multiple instances of the same physics engine (as used by AP) and those that distribute particular aspects of the computational workload of a physics engine. Both techniques will now be discussed.

@@

Techniques that distribute a particular aspect of the workload have not gained any commercial attention recently, unlike techniques that use multiple instances of the same physics engine, which will now be discussed.

@@

\section{Cloud Computing}
In this section a brief overview of cloud computing is given, deployment and service models are discussed. This study's implementation is deployed on the chosen cloud provider, Amazon Web Services (AWS). The latency and packet-loss conditions that affect AP are presented and these values will used to inform the correctness experiments performed on AP.

@@

In streamed gaming, each game instance resides on a single server. There is no technology to balance the real-time requirements of the game across multiple servers. The core problem is that all gaming technology is built and designed for single console/PC install and the greatest bottleneck is the inability to share physics calculations across machines. The technique presented in this study, AP, is intended for use in streamed gaming. AP allows multiple servers to simulate physics, meaning physics simulations are no longer limited in size and complexity to that which can be achieved by only a single server. This allows for game streaming servers to make use of multiple physics servers, the numbers of which can grow and shrink as needed by the game and greatly increase the limit on size and complexity of physics within the game.

\section{Background Summary}
%In physics simulations, there must be an object present in the solving phase for it to be considered in the overall solution of the scenario. A non-migratory approach requires a "ghost" representation of the object in a remote server to enable interaction. Given the calculations and discretisation steps, a "ghost" object takes up just as many resources in deriving a solution as a real object. This rules out the non-migratory approach for our problem. This leaves migratory.
%
%Migratory approaches are concerned with managing the network traffic for those entities that could possibly exist on two servers but can only be solved on one. Such objects need to be placed with an owning server while minimising the effect of thrashing (where an algorithm frequently transfers objects between servers).

For the remainder of this research, we describe our approach to solving this algorithmically, present how a working implementation was achieved, and present results evidencing our work. This is the first presentation of literature that can demonstrate real-time scalable server-side physics modelling and is a significant contribution to reducing the cost of commercialised streamed gaming.
